---
title: "lab_5"
author: "Vienna Saccomanno"
date: "2/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab Week 5 Objectives:

- Intro to time series data class (ts)
- Graphing ts data
- Decomposing ts data
- Exploring autocorrelation
- Holt Winters and ARIMA preview

Required packages:

- tidyverse
- tseries
- forecast

a. Attach packages 

```{r}
library(tidyverse)
library(tseries)
library(forecast)

```


```{r}
energy <- read_csv("energy.csv")
```

- U.S. Residential Energy Consumption (Jan 1973 - Oct 2017)
- Trillion BTU
- US Energy Information Administration (data.gov)

c. Convert to ts data using information in the res_total column from "energy"
```{r}
res_ts<- ts(energy$res_total, frequency = 12, start=c(1973,1)) #January 1973

res_ts
plot(res_ts) #exploratory plot

#Observations:
#-there is an increasing trend
#-These data look additive

```

d. decompose our times series

```{r}

res_dc<- decompose(res_ts)
plot(res_dc)

#Observations: moving average/trend is increasing, seasonality is a pretty big contributor, residuals don't seem to have a pattern
```

e. other cool ways to presen time series data
```{r}
monthplot(res_ts)
#breaks down within a month how values are changing, over all of the years recorded. Dramaticiincrease in residential energy use in summer months
```


```{r}
ggseasonplot(res_ts)
#We see increasing energy use over time

```

f. simple moving averages
-use forecast::ma()
```{r}

sma_res<- ma(res_ts, order=5)
plot(res_ts) #smoothed moving average with window size of x. Changing the window alters how the data appears. Best option is to show the original data + smoothed data
lines(sma_res, col="red")

```

g. Plot ACF for autocorrelation
```{r}
res_acf<-acf(res_ts)

ggtsdisplay(res_ts) #original data + ACF
```

h. Forcasting using exponential smoothing
-holtwinters: level, trend and seasonality
```{r}

res_hw<-HoltWinters(res_ts)
res_hw
#Smoothing parameters:
 #alpha: 0.3258894
 #beta : 0.001104119
 #gamma: 0.3005416

plot(res_hw) #original data with prediction on top to check model accuracy (not yet forecasting)


```

```{r}
#Future forecasting with HW

res_forecast<-forecast(res_hw, h= 60) #forecasting for 5 years, 60 mo

plot(res_forecast)
#flatening trend because weighting more recent years


```
#check distribution of the residuals

```{r}
#forevery month, find difference between actual data and forecast

hist(res_forecast$residuals) #look pretty normal!
```

ARIMA: autoregressive intergrated moving avergare (Arima) using auto.arima for p,d,q
-moving average + trend
```{r}
#Need to find p,d, q

res_pdq<-auto.arima(res_ts)
res_pdq

#ARIMA(1,0,2)(0,1,1) = seasonal and non-seasonal
```

Fit ARIMA model using pdq inputs
```{r}

res_arima<-arima(res_ts, order=c(0,1,2), seasonal = list(order=c(0,1,1)))

#Check out the residuals
par(mfrow = c(1,2))
hist(res_arima$residuals)
qqnorm(res_arima$residuals) #no major concerns
```

```{r}
forecast_res <- forecast(res_arima, h = 72)
plot(forecast_res) #6 years
```

```{r}
#ggplot of the outpus

res_df<-data.frame(forecast_res)
months_seq<- seq(1,72)

res_df_2<-data.frame(months_seq, res_df)


ggplot(res_df, aes(x=months_seq, y=Point.Forecast))+
  geom_line()
```

